Unifying Lower Bounds on Prediction Dimension of Consistent Convex Surrogates

### 2021-01-26 colt-21/cvx-flats.tex
Distributions and consistency notes

If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the $\ell$-regret makes the two definitions are equivalent. % for the appropriate choice of $\propdis$.
% when using target regret of $r$ (compared to an optimal prediction for $p$) as $\propdis(r,p)$.
\begin{lemma}\label{lem:consistent-loss-implies-prop}
	Given a surrogate loss $L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
  $\mu(r,p) := R_\ell(r,p)$.
  Then
	$(L, \psi)$ is $\D$-consistent with respect to $\ell$ if and only if $(L,\psi)$ is $(\propdis, \D)$-consistent with respect to $\gamma := \prop[\simplex]{\ell}$. \jessie{We can restrict to $\P$ right? Is $\P \supseteq \cup_{x \in \X} D_x$ required?}
  \btw{Note: don't actually need $\gamma$ nonempty here.}
\end{lemma}
\jessie{Defer proofs to appendix from this section}
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff \exploss{\ell}{r}{p} = \inf_{r' \in \R} \exploss{\ell}{r'}{p} \iff r \in \gamma(p)$.
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
  Rewriting the right-hand side of Definition~\ref{def:consistent-ell},
        \begin{align}
	  &\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf\nolimits_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	  &\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	  &\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
        Therefore, $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),Y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
\end{proof}

@@

\begin{theorem}\label{thm:consistent-implies-indir-elic}
	For a convex surrogate $L \in \L$, if the pair $(L, \psi)$ is consistent with respect to a property $\gamma$ or a loss $\ell$ eliciting $\gamma$, then $(L, \psi)$ indirectly elicits $\gamma$.\jessiet{How do we want to handle consistency here- we want it to be specific to the distribution}\jessie{Require $\P \subseteq \cup_{D \in \D}\cup_x D_x$ and convex?  }
\end{theorem}
\begin{proof}
%  \raft{Maybe ref the consistency definition here explicitly, at the top and the end}
	By Lemma~\ref{lem:consistent-loss-implies-prop}, it suffices to show the result for consistency with respect to a property $\gamma$, setting $\gamma := \prop{\ell}$ if $\ell$ is given instead.
	We show the contrapositive; suppose $(L, \psi)$ does not indirectly elicit $\gamma$, meaning we have some $p \in \P$ so that $u \in \Gamma(p)$ but $\psi(u) \not \in \gamma(p)$, where $\Gamma := \prop{L}$.
	Observe that we use the fact $\Gamma(p) \neq \emptyset$.
	Consider the constant sequence $\{f_m\}$ with $f_m(x) = u$ for all $m,x$, and take $D$ with full support on some $x \in \X$, and let $D_x = p$.
  Since $u \in \Gamma(p)$, we observe $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$; in particular $\E_D L(f_m(X), Y) \to \inf_f L(f(X),Y)$.
%  \raft{Maybe clearer to first say $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$, and then the easy implication of the convergence \jessie{Done}}
	However, we have $\E_X \propdis(\psi \circ f_m(X), D_X) = \propdis(\psi(u_m), p) = \propdis(\psi(u), p) \neq 0$, since $\psi(u) \not \in \gamma(p)$.
%  \raft{Almost -- just clarify that this is the expectation over $D$ too by construction \jessie{Kinda done?}}
	Thus, we observe $(L, \psi)$ is not consistent with respect to $\gamma$ (Definition~\ref{def:consistent-prop}).
\end{proof}

@@

\begin{corollary}\label{cor:Pcodim-flat-single-val-prop}
Let target property $\gamma:\P \toto \R$ and $d\in\mathbb N$ be given.
Let $p \in \P$ with $|\gamma(p)| = 1$, and take $\gamma(p) = \{r\}$.
If there is no $d$-flat $F$ with $p \in F \subseteq \gamma_r$, then no $L\in\L_d$ indirectly elicits $\gamma$\raft{To discuss next meeting: maybe it's better to say $\eliccvx(\gamma) > d$ instead of the indirectly elicits bit.  Both here and the next Cor.}, and in particular, there is no convex surrogate $L \in \L_d$ consistent with respect to $\gamma$.
%  %commented out 20 Jan 2021 8:37am - Jessie.  switching first statement to be the contrapositive
%  Suppose we are given a target property $\gamma:\P \toto \R$, where $\P\subseteq\simplex$, indirectly elicited by a convex loss $L \in \L_d$.
%  For all $p \in \P$ with $|\gamma(p)| = 1$, and $\gamma(p) = \{r\}$, there is some flat $F$ of \Pcodimension at most $d$ such that $p \in F \subseteq \gamma_r$.
%  In particular, if there is a $p \in \P$ such that $\gamma(p) = \{r\}$ such that there is no flat $F$ of \Pcodimension at most $d$ such that $p \in F \subseteq \gamma_r$, then there is no convex surrogate $L \in \L_d$ consistent with respect to $\gamma$.
\end{corollary}
\begin{proof}
	Let $(L, \psi)$ indirectly elicit $\gamma$ and the convex function $L$ and elicit $\Gamma$.
	As $\Gamma$ is non-empty, there is some $u \in \Gamma(p)$.
	Since $\gamma$ is single-valued at $p$, we have $r = \psi(u)$; by Lemma~\ref{thm:convex-flats-inf-dim}, we know there is a flat $F = \zeros{V_{u,p}}$ so that $p \in F \subseteq \Gamma_u$.
	Moreover, by construction, we know it is a $d$-flat.
	By definition of indirect elicitation, we additionally have $\Gamma_u \subseteq \gamma_r$.
	Thus, we have $p \in F \subseteq \gamma_r$.

@@

\begin{restatable}[\cite{ramaswamy2016convex} Theorem 18]{corollary}{hariresult}\label{cor:fsd-bound}
	Let $\ell:\R \times \Y \to \reals$ be a discrete loss eliciting $\gamma:\simplex \toto \R$, and let $L \in \L_d$ be a minimizable consistent convex surrogate for $\ell$.
	Then for all $p \in \simplex$ and $r \in \gamma(p)$,
	\begin{equation}
	  d \geq \|p\|_0 - \dim(\Sc_{\gamma_r}(p)) - 1~.~
	\end{equation}
%% Bo: old version, Jun 4 7pm
%%	Suppose we are given a discrete loss $\ell:\R \to \reals^\Y_+$ eliciting $\gamma$.
%%	Take $p \in \simplex$ and $r \in \R$ such that $p \in \gamma_r$.
%%	\begin{equation}
%%	\ccdim(\ell) \geq \|p\|_0 - \dim(\Sc_{\gamma_r}(p)) - 1~.~
%%	\end{equation}\jessiet{TODO: add minnability\bo{Tried!}}
\end{restatable}
\begin{proof}[Proof sketch]
	%\btw{Bo: rearranged the statements here a little, I felt the previous version required the reader to keep a bit more in their head.}
	\jessie{Redid this citing Lemma~\ref{lem:finite-relint-dim} instead}
	First, take $p \in \relint{\simplex}$ for intuition, so $\|p\|_0 = n$.
	Lemma~\ref{thm:convex-flats-inf-dim} guarantees the existence of a $d$-flat $F$ with $p \in F \subseteq \gamma_r$.
	Moreover, since $p \in \relint{\simplex}$, the flat $F$ satisfies the requirements of Lemma~\ref{lem:finite-relint-dim}, and the bound $\dim(\Sc_{\gamma_r}(p)) \geq \spn(F - \{p\}) \geq n - d- 1$ is given by chaining Lemma~\ref{lem:finite-relint-dim} (ii) and Lemma~\ref{lem:Fp-subset-feasible-subset}.
%	By Lemma~\ref{lem:Pcodim-codim-finite}, this implies the existence of a matrix $W$ such that $p \in \{q \in \simplex : Wq = \vec 0 \} = F \subseteq \gamma_r$ and $\codim(W) \leq d$.
%	%dimension at least $n-d-1$ with $p \in F$ and $F \cap \simplex \subseteq \gamma_r$. 
%	For $p \in \relint{\simplex}$, this implies that $\Sc_{\gamma_r}(p) \supseteq F - p$.
%	Moreover, this implies $\dim(\Sc_{\gamma_r}(p)) \geq \codim(W) = n - d - 1$, completing the proof for $p \in \relint{\simplex}$.

>>>>

If $\gamma$ is a property elicited by a target loss $\ell$, then taking $\propdis$ to be the $\ell$-regret makes the two definitions are equivalent. % for the appropriate choice of $\propdis$.
% when using target regret of $r$ (compared to an optimal prediction for $p$) as $\propdis(r,p)$.
\begin{lemma}\label{lem:consistent-loss-implies-prop}
	Given a surrogate loss $L$, link $\psi$, and target loss $\ell$, set
	% Take $\propdis: \R \times \simplex \to \reals_+$ with 
  $\mu(r,p) := R_\ell(r,p)$.
  Then
	$(L, \psi)$ is $\D$-consistent with respect to $\ell$ if and only if $(L,\psi)$ is $(\propdis, \D)$-consistent with respect to $\gamma := \prop[\simplex]{\ell}$. \jessie{We can restrict to $\P$ right? Is $\P \supseteq \cup_{x \in \X} D_x$ required? I think here, having the property over the simplex is fine, but not sure how that propagates to the theorem 9.}
  \btw{Note: don't actually need $\gamma$ nonempty here.}
\end{lemma}
\jessie{Defer proofs to appendix from this section}
\begin{proof}
	First, observe that $\propdis(r,p) = 0 \iff \exploss{\ell}{r}{p} = \inf_{r' \in \R} \exploss{\ell}{r'}{p} \iff r \in \gamma(p)$.
	Now suppose $(L, \psi)$ are consistent with respect to $\ell$, and take any sequence $\{f_m\}$ of measurable hypotheses.
  Rewriting the right-hand side of Definition~\ref{def:consistent-ell},
        \begin{align}
	  &\; \E_D \ell(\psi \circ f_m(X), Y)\to \inf\nolimits_f \E_D \ell(\psi \circ f(X), Y)   \label{eqn:cons-loss-cond} \\
	  &\iff \E_X R_\ell(\psi \circ f_m(X), D_X) \to 0                               \nonumber  \\
	  &\iff \E_X \propdis(\psi \circ f_m(X), D_X) \to 0~.~                          \label{eqn:cons-prop-cond}
	\end{align}
        Therefore, $\mathbb{E}_D L(f_m(X),Y) \to \inf_f \mathbb{E}_D L(f(X),Y)$ implies (\ref{eqn:cons-loss-cond}) if and only if it implies (\ref{eqn:cons-prop-cond}).
\end{proof}

@@

\begin{theorem}\label{thm:consistent-implies-indir-elic}
	For a convex surrogate $L \in \L$, if the pair $(L, \psi)$ is $\D$-consistent with respect to a property $\gamma$ or a loss $\ell$ eliciting $\gamma$, then $(L, \psi)$ indirectly elicits $\gamma$.\jessiet{How do we want to handle consistency here- we want it to be specific to the distribution}\jessie{Require $\P \subseteq \cup_{D \in \D}\cup_x D_x$ and convex?  - This requirement works, but may be strictly tighter than we need.  With our assumptions on the loss (namely, the expected loss is finite everywhere for all $p$), we should be able to apply Fubini-Tonelli, so that $\P \subseteq \cup_D \cup_x \D_x$ is sufficient, as any $\D_X$ will be covered by this since the joint expectation is the same as $\E_X \E_Y L(u;D_X)$}\jessiet{Need $\X$ and $\Y$ to be $\sigma$-finite, which I think is immediate from them being probability measures}
\end{theorem}
\begin{proof}
%  \raft{Maybe ref the consistency definition here explicitly, at the top and the end}
	By Lemma~\ref{lem:consistent-loss-implies-prop}, it suffices to show the result for consistency with respect to a property $\gamma$, setting $\gamma := \prop{\ell}$ if $\ell$ is given instead.
	We show the contrapositive; suppose $(L, \psi)$ does not indirectly elicit $\gamma$, meaning we have some $p \in \P$ so that $u \in \Gamma(p)$ but $\psi(u) \not \in \gamma(p)$, where $\Gamma := \prop{L}$.
	Observe that we use the fact $\Gamma(p) \neq \emptyset$.
	Consider the constant sequence $\{f_m\}$ with $f_m(x) = u$ for all $m,x$, and take $D$ with full support on some $x \in \X$, and let $D_x = p$.
  Since $u \in \Gamma(p)$, we observe $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$; in particular $\E_D L(f_m(X), Y) \to \inf_f L(f(X),Y)$.
%  \raft{Maybe clearer to first say $\E_D L(f_m(X), Y) = \inf_f L(f(X),Y)$ for all $m$, and then the easy implication of the convergence \jessie{Done}}
	However, we have $\E_X \propdis(\psi \circ f_m(X), D_X) = \propdis(\psi(u_m), p) = \propdis(\psi(u), p) \neq 0$, since $\psi(u) \not \in \gamma(p)$.
%  \raft{Almost -- just clarify that this is the expectation over $D$ too by construction \jessie{Kinda done?}}
	Thus, we observe $(L, \psi)$ is not consistent with respect to $\gamma$ (Definition~\ref{def:consistent-prop}).
\end{proof}

@@

\jessie{Here, we need to mention $\D$-consistency in the ``in particular''s.  And specify $\P \subseteq \cup_{D \in \D} \cup_x D_x$ or whatever condition we settle on.}
\begin{corollary}\label{cor:Pcodim-flat-single-val-prop}
Let target property $\gamma:\P \toto \R$ and $d\in\mathbb N$ be given.
Let $p \in \P$ with $|\gamma(p)| = 1$, and take $\gamma(p) = \{r\}$.
If there is no $d$-flat $F$ with $p \in F \subseteq \gamma_r$, then no $L\in\L_d$ indirectly elicits $\gamma$\raft{To discuss next meeting: maybe it's better to say $\eliccvx(\gamma) > d$ instead of the indirectly elicits bit.  Both here and the next Cor.}, and in particular, there is no convex surrogate $L \in \L_d$ consistent with respect to $\gamma$.
%  %commented out 20 Jan 2021 8:37am - Jessie.  switching first statement to be the contrapositive
%  Suppose we are given a target property $\gamma:\P \toto \R$, where $\P\subseteq\simplex$, indirectly elicited by a convex loss $L \in \L_d$.
%  For all $p \in \P$ with $|\gamma(p)| = 1$, and $\gamma(p) = \{r\}$, there is some flat $F$ of \Pcodimension at most $d$ such that $p \in F \subseteq \gamma_r$.
%  In particular, if there is a $p \in \P$ such that $\gamma(p) = \{r\}$ such that there is no flat $F$ of \Pcodimension at most $d$ such that $p \in F \subseteq \gamma_r$, then there is no convex surrogate $L \in \L_d$ consistent with respect to $\gamma$.
\end{corollary}
\begin{proof}
	Let $(L, \psi)$ indirectly elicit $\gamma$ and the convex function $L$ and elicit $\Gamma$.
	As $\Gamma$ is non-empty, there is some $u \in \Gamma(p)$.
	Since $\gamma$ is single-valued at $p$, we have $r = \psi(u)$; by Lemma~\ref{thm:convex-flats-inf-dim}, we know there is a flat $F = \zeros{V_{u,p}}$ so that $p \in F \subseteq \Gamma_u$.
	Moreover, by construction, we know it is a $d$-flat.
	By definition of indirect elicitation, we additionally have $\Gamma_u \subseteq \gamma_r$.
	Thus, we have $p \in F \subseteq \gamma_r$.

@@

\begin{restatable}[\cite{ramaswamy2016convex} Theorem 18]{corollary}{hariresult}\label{cor:fsd-bound}
	Let $\ell:\R \times \Y \to \reals$ be a discrete loss eliciting $\gamma:\simplex \toto \R$, and let $L \in \L_d$ be a minimizable consistent convex surrogate for $\ell$.
	Then for all $p \in \simplex$ and $r \in \gamma(p)$,
	\begin{equation}
	  d \geq \|p\|_0 - \dim(\Sc_{\gamma_r}(p)) - 1~.~
	\end{equation}
%% Bo: old version, Jun 4 7pm
%%	Suppose we are given a discrete loss $\ell:\R \to \reals^\Y_+$ eliciting $\gamma$.
%%	Take $p \in \simplex$ and $r \in \R$ such that $p \in \gamma_r$.
%%	\begin{equation}
%%	\ccdim(\ell) \geq \|p\|_0 - \dim(\Sc_{\gamma_r}(p)) - 1~.~
%%	\end{equation}\jessiet{TODO: add minnability\bo{Tried!}}
\end{restatable}
\begin{proof}[Proof sketch]
	%\btw{Bo: rearranged the statements here a little, I felt the previous version required the reader to keep a bit more in their head.}
	\jessie{Redid this citing Lemma~\ref{lem:finite-relint-dim} instead}
	First, take $p \in \relint{\simplex}$ for intuition, so $\|p\|_0 = n$.
	Lemma~\ref{thm:convex-flats-inf-dim} guarantees the existence of a $d$-flat $F$ with $p \in F \subseteq \gamma_r$.
	Moreover, since $p \in \relint{\simplex}$, the flat $F$ satisfies the requirements of Lemma~\ref{lem:finite-relint-dim}, and the bound $\dim(\Sc_{\gamma_r}(p)) \geq \Scr_F(p) \geq n - d- 1$ is given by chaining Lemma~\ref{lem:finite-relint-dim} (ii) and Lemma~\ref{lem:Fp-subset-feasible-subset}.
%	By Lemma~\ref{lem:Pcodim-codim-finite}, this implies the existence of a matrix $W$ such that $p \in \{q \in \simplex : Wq = \vec 0 \} = F \subseteq \gamma_r$ and $\codim(W) \leq d$.
%	%dimension at least $n-d-1$ with $p \in F$ and $F \cap \simplex \subseteq \gamma_r$. 
%	For $p \in \relint{\simplex}$, this implies that $\Sc_{\gamma_r}(p) \supseteq F - p$.
%	Moreover, this implies $\dim(\Sc_{\gamma_r}(p)) \geq \codim(W) = n - d - 1$, completing the proof for $p \in \relint{\simplex}$.
